|

### 2020-12-10 01-Chp-Introduction.Rmd
Fix fancy headers in template file and update other files
End of day: -700

# Introduction {#intro-intro}
\minitoc <!-- this will include a mini table of contents-->

@@

```{r load_helper_scripts, include=FALSE}
source("R/citations.R")
```

@@

@@

* __Chapter \@ref(background-heading):__ Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis. 
* __Chapter \@ref(sys-rev-heading)__ This Chapter describes a comprehensive systematic review and meta-analysis of all available evidence on the relationship between blood lipids and lipid 
* __Chapter \@ref(cprd-analysis):__ This Chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database, based in England. 
* __Chapter \@ref():__ This Chapter seeks

* __Appendix \@ref(sys-rev-tools-heading)__ This Chapter introduces new evidence synthesis two tools built in R: `robvis` and `medrixvr`. The motivation for, development and impact of, and future plans for these tools are outlined.

@@

@@

* McGuinness et al.

@@

A lot of packages were used in this project`r gen_rmd_citation()`

* `robvis`: An R package and associated `shiny` web application that allows users to easily visualise the results of risk of bias graphs.

* `medrxvir`: An R package and associated `shiny` web application that allows users to easily search and retrieve bibliographic data from the medRxiv^[https://www.medrxiv.org/] preprint repository.

@@

* Talk at Cochrane on medrxivr
* Presentation to the 
* Webinar to Evidence Synthesis Ireland on Risk of Bias

>>>>

# Introduction {#intro-heading}
\minitoc <!-- this will include a mini table of contents-->

@@

First I provide a background to the central concepts and themes of this thesis, and justify the motivations for this piece of work. I introduce a tool that allows for systematic searching of preprint literature, which is then used in a systematic review of all available evidence on my central research question.

@@

## Theoretical framework
The main theoretical framework used in this thesis is evidence synthesis - the discovery and critical integration of all available evidence on a research question in order to either: a) provide a more definitive answer to that question or; b) highlight gaps in the existing evidence base so that future research address questions that have yet to be answered, or explores the same question in a way that increases our confidence in the result.

All elements of this thesis are based around this framework.

@@

* __Chapter \@ref(background-heading):__ Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis. 
* __Chapter \@ref(sys-rev-tools-heading)__ This CHhapter introduces a new tool, the `medrxivr` R package, which was used to developed to allow for systematic searches of the bioRxiv and medRxiv preprint repositories.
* __Chapter \@ref(sys-rev-heading)__ This Chapter describes a comprehensive systematic review and meta-analysis of all available evidence on the relationship between blood lipids and lipid. 
* __Chapter \@ref(cprd-analysis-heading):__ This Chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database, based in England. 
* __Chapter \@ref(ipd-heading):__ This Chapter describes an individual patient data analysis of several longitudinal cohort studies, to describe the relationship between blood serum lipids and dementia outcomes.

@@

  * McGuinness, Luke A., and Lena Schmidt. "medrxivr: Accessing and searching medRxiv and bioRxiv preprint data in R." Journal of Open Source Software 5.54 (2020): 2651. DOI: [10.21105/joss.02651](https://doi.org/10.21105/joss.02651)

  * McGuinness, Luke A., and Julian PT Higgins. "Risk‐of‐bias VISualization (robvis): An R package and Shiny web app for visualizing risk‐of‐bias assessments." Research Synthesis Methods (2020). DOI: [10.1002/jrsm.1411](https://doi.org/10.1002/jrsm.1411)

@@

  *

@@

* `medrxvir`: An R package and associated `shiny` web application that allows users to easily search and retrieve bibliographic data from the medRxiv^[https://www.medrxiv.org/] and bioRxiv^[https://www.biorxiv.org/] preprint repositories. See Chapter \@ref(sys-rev-tools-heading) for more details. Install a stable version of the package from CRAN with `install.packages("medrxivr")` or alternatively install the development version from GitHub using `devtools::install_github("mcguinlu/medrxivr")`. Link: [https://ropensci.github.io/medrxivr/](https://ropensci.github.io/medrxivr/)

* `robvis`: An R package and associated `shiny` web application that allows users to easily visualize the results of the risk-of-bias assessments performed as part of a systematic review. See Appendix \@ref(appendix-robvis) for more details. Install a stable version of the package from CRAN using `install.packages("robvis")` or alternatively install the development version from GitHub using `devtools::install_github("mcguinlu/robvis")`. Link: [https://mcguinlu.github.io/robvis/](https://mcguinlu.github.io/robvis/)

@@

* Presentation to the 
* Webinar to Evidence Synthesis Ireland on Risk of Bias
* Presentation(s) to the Evidence Synthesis and Meta-Analysis in R Conference (ESMARConf)

### 2020-12-10 03-Chp-Systematic-Review-Tools.Rmd
Update to cite rbiorxiv
Fix fancy headers in template file and update other files
End of day: -700

<!----------------------------------------------------------------------------->
## Development

@@

Early versions of the tool had a reliance on web-scraping data directly from the medRxiv website. Web-scraping is a fragile, or brittle, way to extract data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002;@laprie1992].

However, an Application Programming Interface (API) for the medRxiv and bioRxiv repositories was released in early 2020, allowing for a newer version of the medrxivr package to engage in "fault prevention" and work towards a more robust interface with the available data.

Finally, while initial versions of the tool focused primarily on the medRxiv repository, discovery of some eligible required expansion to the bioRxiv repository, so that both can now be searched from a single tool. However, the introductory examples given below apply are taken from searches performed in the medRxiv repository.

Developed to meet three criteria:

1. reproducible and transparent search functionality, with Boolean operator logic;  
2. support for bulk export of references returned by the search;  
3. automated access full-text records of relevant records.

<!-----------------------------------------------------------------------------> 
## Installation

`medrxivr` has been released to the Comprehensive R Archive Network (CRAN), and can be installed with the following code

To install the stable `medrxivr` from CRAN:

@@

Alternatively, the  install the development version from GitHub:

@@

<!----------------------------------------------------------------------------->
## Importing preprint meta data

@@

<!----------------------------------------------------------------------------->
## Creating a search

@@

<!----------------------------------------------------------------------------->
## Further functionality

<!----------------------------------------------------------------------------->
### Exporting to bibliography

One of the key features of the `medrxivr` is the ability for users to easily export the results of their systematic search to a reference manager. While it is a seemingly simple request, this is is one of the key ways in which `medrxivr` - as it was built with systematic reviews specifically in mind, it is

@@

<!----------------------------------------------------------------------------->
### PDF download

@@

<!----------------------------------------------------------------------------->
### Search reporter

@@

### Data visualisation

@@

<!----------------------------------------------------------------------------->
## Comparison with other search options

During development of the package, there were several alternative considered as part of a audit of existing tools. However, development of a new custom tool was preferred as none met the three criteria required: 1) reproducible and transparent search functionality, with Boolean operator logic; 2) support for bulk export of references returned by the search; 3) automated access full-text records of relevant records.

Some previous tools exist while allowed for robust searching of preprint repositories, such as `search.bioPreprint` [@iwema2016]- however, these tools were aimed more at those looking to keep up to date with recent developments rather than systematically assess the entirety . As such, they did not support needed functionality such as bulk export of records matching a search term, unlimited return of records (c.f. `search.bioPreprint` which is limited to the most recent 1000 records matching a user search).

The `medrxvir` package was written in R using RStudio, and followed development best practices, including complete and information documentation, a robust unit testing framework (99% of all code lines within the package are formally tested under this framework) across multiple platforms including Windows, MacOS, and Linux, and in-depth code review by two experienced reviewers.

The medRxiv snapshot is still taken every morning using GitHub Actions, an automated system for repetitive tasks.

An automatically generated documentation website is also generated on each update to the code-base ^[https://docs.ropensci.org/medrxivr/]

@@

it is too early to see if preprint reps

There is the potential that the cross-section of literature posted on medrxiv would be substantially different from the true grey literature - simply lowering the barriers to publication may encourage authors to published "null" results,[CITE] but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[CITE]

As

@@

``` {r, eval = TRUE, echo = FALSE}

t<-rbiorxiv::biorxiv_summary()

```

@@

test

>>>>

(ref:medrxivr-sr-cap) __Role of `medrxivr` in a systematic review workflow:__ `medrxivr` allows for systematic searching of biomedical preprints as part of the initial literature searching. Following title and abstract screening, reviewers can then programmatically retrieve a copy of the PDF of included records to faciliate the full-text screening stage (similar to Endnote's "Find Full Text" feature).

(ref:medrxivr-sr-scap) Role of `medrxivr` in a systematic review workflow

```{r medrxivr-sr, echo = FALSE, results="asis", fig.cap='(ref:medrxivr-sr-cap)', out.width='100%', fig.scap='(ref:medrxivr-sr-scap)'}
knitr::include_graphics(file.path("figures/UoBcrest.pdf"))
```

## Development

Development of the package followed the

### Success crtieria

Defining success criteria

The tool was developed to meet three success criteria,[@wateridge1995] created both from the list of functionality needed to perform the systematic searches as part of the review in Chapter \@ref(sys-rev-heading) and from an informal survey of colleagues and the evidence synthesis and health librarians on Twitter. These criteria were as follows:

1. reliable, reproducible and transparent search functionality, allowing for similar Boolean (AND/OR/NOT) operator logic as that available from other literature search platforms such as OvidSP;  
2. support for bulk export of references returned by the search to a file type that can be imported into a reference manager (e.g., .bib  or .ris);  
3. automated retrieval of the full-text PDFs of relevant records.

### Early version

@@

Early versions of the tool had a reliance on web-scraping data directly from the medRxiv website. Web-scraping is a fragile mechanism for extracting data, as it is entirely dependent on consistent website design and underlying code structure remaining unchanged.[@shaw2002;@laprie1992]. In the case of early versions of `medrxivr`, constant maintenance work was required to ensure the web-scraping script performed as expected, as the medRxiv website was regularly updated.

However, an Application Programming Interface (API) for the medRxiv and bioRxiv repositories was released in early 2020, allowing for a newer version of the medrxivr package to engage in "fault prevention" and work towards a more robust interface with the available data.[@laprie1992]

Finally, while initial versions of the tool focused primarily on the medRxiv repository, discovery of some eligible papers for the review detailed in Chapter \@ref(sys-rev-heading) required expansion to allow for similar searching of the bioRxiv repository.

The `medrxvir` package was written in R using RStudio, and followed development best-practice, including complete and information documentation, a robust unit testing framework (99% of all code lines within the package are formally tested under this framework) across multiple platforms including Windows, MacOS, and Linux, and in-depth code review by two experienced, independent reviewers.

An automatically generated documentation website is also generated on each update to the code-base. ^[https://docs.ropensci.org/medrxivr/]

<!-----------------------------------------------------------------------------> 
## Usage

## Installation

`medrxivr` has been released to the Comprehensive R Archive Network (CRAN), and can be installed with the following code:

@@

Alternatively, the development version of the pacakage can be installed from GitHub:

@@

<!----------------------------------------------------------------------------->
### Importing preprint metadata

@@

<!----------------------------------------------------------------------------->
### Performing a search

@@

<!----------------------------------------------------------------------------->
### Exporting to a bibliography file

In line with the second success criteria, one of the key features of the `medrxivr` is the ability for users to easily export the results of their systematic search to a reference manager. While it is a seemingly simple request, this is is one of the key ways in which `medrxivr` - as it was built with systematic reviews specifically in mind, it is

@@

<!----------------------------------------------------------------------------->
### Downloading the PDFs of relevant records

@@

<!----------------------------------------------------------------------------->
### Reporting your results

@@

### Visualising your results

@@

<!----------------------------------------------------------------------------->
### Comparison with other bioRxiv/medRxiv interfaces

Prior development of `medrxivr`, several alternative search tools were considered as part of a audit of existing tools for accessing the metadata of records stored on the medRxiv and bioRxiv servers.

Two key examples are useful to consider to highlight the additional functionality that `medrxivr` contributes.

The first, a platform called Rxivist [@abdill2019], Similar to `medrxivr`, this tool previously scrapped data straight from the bioRxiv. (at the time of development, Rxvist , but as of November 2020, Rxivist also provides access to

However, this platform is focused more on exploring the number of times a preprint has been downloaded and/r shared on Twitter, to allow researchers to find the most popular papers related to their topic. The search interface^[Available at https://rxivist.org/] does not allow for either complex search strategies or 
The second `search.bioPreprint`, allows users to search for terms across a range of preprint servers, including medRxiv, bioRxiv, but also publishing venues such as F1000Research.  [@iwema2016]. However, this tool is aimed more at those looking to keep up to date with recent developments rather than systematically assessing the entirety of the available literature, and as such, only returns the most recent `r comma(1000)` records by publication date.

Finally, neither tool provides an easy way to programmatically download a copy of the PDF of relevant records.

. As such, this tool does not meet the second success did not meet the success criteria laid out in the introduction to this Chapter.

@@

There is the potential that the cross-section of literature posted on medrxiv would be substantially different from the true grey literature - simply lowering the barriers to publication may encourage authors to published "null" results,[CITE] but due to the effort involved in writing up a distributable manuscript, it is unlikely to completely address the "file drawer" effect.[CITE] It is likely too early, (and too difficult) to tell whether. **The additional contribution of searching these repositories to the systematic review element of this thesis is explored in Chapter \@ref(sys-rev-heading)**

@@

@@

### 2020-12-10 04-Chp-Systematic-Review.Rmd
End of day: -700

# Systematic Review {#sys-rev-heading}
\minitoc <!-- this will include a mini table of contents-->

@@

@@

#### Inclusion
We will seek studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia/MCI. Eligible study designs include randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

@@

#### Exclusion
Case-control studies, cross-sectional studies, qualitative studies, case reports/series and narrative reviews will be excluded. Studies which present no evidence of attempting to exclude prevalent cases from their analyses will also be excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/MCI/dementia) will be excluded. Conference abstracts with no corresponding full-text publication will be examined, and we will contact authors to obtain information on the study’s status. Studies that are reported in insufficient detail (e.g. only in conference abstracts, new, letters, editorials and opinion) will be excluded. Previous systematic reviews will not be eligible, but their reference lists will be screened to identify any potentially relevant articles. Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a “multi-domain intervention” where the lipid-regulating agent is included in each arms (e.g. for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e. below the age of 50) will be excluded.

@@

Harmonization of cholesterol measures across studies was performed, as different studies used different methods to quantify exposure, including comparing differing risks in the highest vs lowest quartiles of a lipid, using a binary classification of patients into a hypercholesterolaemia or not, categorising lipid levels into high, middle, and low groups according to study-defined criteria, and simply treating the exposure as a continuous variable.

@@

Risk of bias assessment was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomised controlled trials were assessed using the RoB2 tool,[@sterne2019], non-randomised studies of interventions were assessed using the ROBINS-I tool,[@sterne2016] and non-randomised studies of exposures were assessed using the ROBINS-E tool.[TBC]

At present, no risk of bias assessment tool for Mendelian randomisation studies is available. Bias in thesee studies was assessed with the help of an expert panel [TBC]

Data will be visualized using a paired forest and risk of bias plot

@@

#### Inter-rater reliability

@@

<!----------------------------------------------------------------------------->
``` {r chunkname_setup, results="hide", message=FALSE}
```

```{r chunkname_table, message=FALSE, results="asis"}
```
<!----------------------------------------------------------------------------->

#### PRISMA flowchart

Following de-duplication, the titles and abstracts of 16109 records were assessed for eligibility. 387 were deemed potentially eligible and the full text records for these were requested and screened. [**CROSS-REF to PRISMA flow here**]

A breakdown of records by

## Included studies

@@

## Publication bias

@@

### Triangualtion across evidence sources

@@

A key limitiation for other types of dementia, in particularly vascular, is that there has yet to be a GWAS identifying relevant SNPS that could then be used in a MR study with SNPS for lipids to estimates the causal effect of lipids on vascular dementia [**Is this true?**] This rules out the use

@@

Of note, as part of the review, we identified serveral previous systematic reviews of this topic. However, this review is the firest to

Consideration of the potential impact of the magnitude and direct of unaddressed confounders/bias is not a major stretch from what is already happening in the assessment of the quality of evidence (GRADE) framework. Wihtin GRADE the overall quality of evidence can be upgraded when there is deemed to be **will need to confirm this.l** [@guyatt2011]

>>>>

# Systematic review and triangulation of available evidence {#sys-rev-heading}
\minitoc <!-- this will include a mini table of contents-->

@@

## To Do

* Examine how many times a primary study was included in a review  
* Look at wether searching preprints made any difference to my results

@@

#### Inclusion criteria
We will seek studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia/MCI. Eligible study designs include randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

@@

#### Exclusion criteria
Case-control studies, cross-sectional studies, qualitative studies, case reports/series and narrative reviews will be excluded. Studies which present no evidence of attempting to exclude prevalent cases from their analyses will also be excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/MCI/dementia) will be excluded. Conference abstracts with no corresponding full-text publication will be examined, and we will contact authors to obtain information on the study’s status. Studies that are reported in insufficient detail (e.g. only in conference abstracts, new, letters, editorials and opinion) will be excluded. Previous systematic reviews will not be eligible, but their reference lists will be screened to identify any potentially relevant articles. Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a “multi-domain intervention” where the lipid-regulating agent is included in each arms (e.g. for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e. below the age of 50) will be excluded.

@@

Harmonization of cholesterol measures across studies was performed, as different studies used different methods to quantify exposure, including comparing differing risks in the highest vs lowest quartiles of a lipid, using a binary classification of patients into a hypercholesterolaemia or not, categorizing lipid levels into high, middle, and low groups according to study-defined criteria, and simply treating the exposure as a continuous variable.

@@

Risk of bias assessment was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomized controlled trials were assessed using the RoB2 tool,[@sterne2019], non-randomized studies of interventions were assessed using the ROBINS-I tool,[@sterne2016] and non-randomized studies of exposures were assessed using the ROBINS-E tool.[CITE]

At present, no risk of bias assessment tool for Mendelian randomization studies is available. Bias in these studies was assessed with the help of an expert panel drawn from my supervisors and other external invited researchers.[CITE]

Data will be visualized using a paired forest and risk of bias plot, created using the `robvis` tool. This tool was developed as part of this thesis to aid in creating standard risk-of-bias figures, and is summarized in Appendix \@ref(appendix-robvis).[@mcguinness2019]

@@

@@

<!----------------------------------------------------------------------------->
``` {r prisma-flow-setup, results="hide", message=FALSE}
```

(ref:prisma-flow-cap) __PRISMA Flowchart:__ Overview of how records moved through the systematic review process

(ref:prisma-flow-scap) PRISMA Flowchart

```{r prisma-flow-fig, echo = FALSE, results="asis", fig.cap='(ref:prisma-flow-cap)', out.width='100%', fig.scap='(ref:prisma-flow-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```
<!----------------------------------------------------------------------------->

Following de-duplication, the titles and abstracts of 16109 records were assessed for eligibility. 387 were deemed potentially eligible and the full text records for these were requested and screened. \@ref(fig:prisma-flow-fig)

### Characteristics of included studies

@@

### Risk of bias {#risk-of-bias-subheading}

### Publication bias

@@

## Triangualtion across evidence sources

@@

A key limitation for other types of dementia, in particularly vascular, is that there has yet to be a GWAS identifying relevant SNPS that could then be used in a MR study with SNPS for lipids to estimates the causal effect of lipids on vascular dementia [**Is this true?**] This rules out the use. In addition, there was limited literature available - as discussed in CHapter \@ref()

Consideration of the potential impact of the magnitude and direct of residual confounders/bias is not a major stretch from what is already happening in the assessment of the quality of evidence (GRADE) framework. Within GRADE, the overall quality of evidence can be upgraded when there is deemed to be unmeasured or residual confounding variables which reduce the. For example, if the propensity to treatment is related to comorbidity burden, but those on treatment still have better outcomes then those on control, it is likely that the true effect of the intervention is being underestimated. [@guyatt2011]

Without our framework and as part of the risk of bias assessments reported in \@ref(risk-of-bias-subheading), we test  test

@@

### Comparison with other reviews

Of note, as part of the review, we identified several previous systematic reviews of this topic.[CITE] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies, and explore the heterogeneity of results across different levels of risk of bias levels.

Some previous reviews did assess risk of bias, but used non-domain based assessment tools. Newcastle-Ottowa scale,

__The duplication of work across reviews (including, ironically, by this review) is substantial. In Section XXXX, I demonstrated that the each primary study included in this review was also included in other review so this topic, but that not all studies were included in all reviews. However, by creating a comprehensive review that attempted to draw together all available evidence from across the full range of study types.__

### Strenghts and limitations

### 2020-12-10 front-and-back-matter/99-appendices.Rmd
End of day: -700

# By Chapter

## Chapter 1

## Chapter 2

# Other Appendix

<!----------------------------------------------------------------------------->
## Producing risk-of-bias visualisation with `robvis`

>>>>

# By Chapter {#chapter-appendix-heading}

## Chapter 1 {#appendix-chapter-1}

## Chapter 2 {#appendix-chapter-2}

<!----------------------------------------------------------------------------->
# Other Appendix {#other-appendix-heading}

## Software used to create this thesis

This thesis was written in RMarkdown.
Several R packages were used as part of this project.`r gen_rmd_citation()`

## Producing risk-of-bias visualisation with `robvis` {#appendix-robvis}
